---
title: "PCA-multivariate"
format: pdf
---

## 

Load the reduced to the top 9 species in the Gulf data
```{r load data, echo=FALSE}
# Load required libraries
library(FactoMineR)
library(ggplot2)  # for visualization
# Load the GI data - with inflation adjusted values
top_10_species <- readRDS("~/Resilience/UpdatedData/top_10_data.rds")

```
Brief summary
```{r}
# Display a summary of the GI data
summary(top_10_species)

```

##Top 10 most valuable species
```{r}
head(top_10_species)
names(top_10_species)
#unique_species <- unique(top_10_species$common_name_cleaned)
#print(unique_species)
```
```{r}
library(dplyr)

# Assuming top_10_species is your original dataset
top_10 <- top_10_species %>%
  select(-data_supplier_st, -zipcode, -Inflation_index, -Adjust_factor, -common_name, -species_itis)
summary(top_10)
```

### Aggregate the data for PCA
GOAL: focus on the relationship between a maximum drop in landed values in one year and the increase in landed values in a different species
To do so I do not need info on the city landed - I only need info on the years and various species
Start by aggregating to 

```{r}

library(dplyr)

# Assuming top_10_species is your original dataset
top_10_aggregated <- top_10 %>%
  group_by(landed_year, common_name_cleaned) %>%
  summarise(
    total_live_lbs = sum(live_lbs, na.rm = TRUE),
    total_Adjusted_Val = sum(Adjusted_Val, na.rm = TRUE)
 )



```
##PCA
```{r}


library(tidyr)

# Pivot the data to create a wide-format time series for each year and species
top_10_species_wide <- top_10_aggregated %>%
  pivot_wider(names_from = common_name_cleaned, 
              values_from = c("total_live_lbs", "total_Adjusted_Val"),
              values_fill = 0)

# Assuming top_10_species_wide is your wide-format dataset
# Note: Exclude non-numeric columns (e.g., year and species) from the PCA analysis
pca_result <- prcomp(top_10_species_wide[, -c(1:2)], scale. = TRUE)

# Explore the results
summary(pca_result)

```

The PC1 is strongly associated with the "Blue Crab" species, it means that the variations in the "Blue Crab" data contribute significantly to the overall variability that PC1 is capturing. In other words, when PC1 is high, it indicates a pattern in the data where the "Blue Crab" species tends to have higher values, and when PC1 is low, it suggests lower values for the "Blue Crab."

In practical terms, this can be interpreted as saying that the "Blue Crab" species has the most influence on the dominant patterns or trends represented by PC1. The higher the loading (positive or negative) of a species on PC1, the more it contributes to the overall structure that PC1 is capturing.



```{r}
# Assuming pca_result is your PCA result
loadings <- pca_result$rotation

# Extract loadings for PC1
loadings_PC1 <- loadings[, 1]

# Order the loadings by magnitude
loadings_PC1_ordered <- loadings_PC1[order(abs(loadings_PC1), decreasing = TRUE)]

# Display the top contributing species for PC1
top_species_PC1 <- names(loadings_PC1_ordered)
top_species_PC1

```

Plot
```{r}
# Assuming pca_result is your PCA result
# Assuming species_loadings is a matrix containing the loadings of each species on each PC

# Extracting scores for each observation
scores <- as.data.frame(pca_result$x[, 1:2])

# Extracting loadings for each species
loadings <- as.data.frame(loadings[, 1:2])

# Creating a biplot
biplot <- ggplot() +
  geom_point(data = scores, aes(x = PC1, y = PC2), size = 3) +
  geom_text(data = loadings, aes(x = PC1, y = PC2, label = rownames(loadings)), hjust = 0, vjust = 0) +
  labs(x = "PC1", y = "PC2", title = "Biplot of PCA")

# Creating a biplot with arrows
biplot2 <- ggplot() +
  geom_point(data = scores, aes(x = PC1, y = PC2), size = 3) +
  geom_segment(data = loadings, aes(x = 0, y = 0, xend = PC1, yend = PC2, color = rownames(loadings)),
               arrow = arrow(length = unit(0.03, "npc")), size = 1, alpha = 0.7) +
  scale_color_discrete(name = "Species") +  # Adding legend
  theme(legend.position = "right") +  # Adjusting legend position
  labs(x = "PC1", y = "PC2", title = "Biplot of PCA")

# Display the biplot
print(biplot)


# Display the biplot
print(biplot2)


```

### Filter Species based on loadings
## Interpretation: 

Species with positive loadings on PC1 based on Values are: Snapper, Blue Crab, eastern Oyster, Stone crab and Groupeer
Species with positive loadings on PC1 based on Pounds are: Snapper, White Shrimp, Eastern Oyster and Grouper

==> these are the species that contribute to an increase in overall variance. Higher values of Snapper, Blue Crab, oyster and grouper are associated with a positive direction of PC1. 

Species with negative loadings on PC1 based on Values: White Shrimp, Pink Shrmp, Brown Shrimp, Spiny Lobster
Species with negative Loadings on PC1 based on pounds: Pink Shrimp, Brown Shrimp, Spiny Lobster, Stone Crab



```{r}
# Assuming pca_result is your PCA result
loadings <- pca_result$rotation

# Extract loadings for PC1 and PC2
loadings_PC1 <- loadings[, 1]
loadings_PC2 <- loadings[, 2]

# Create a data frame with species and their loadings on PC1 and PC2
loadings_data <- data.frame(species = rownames(loadings), PC1 = loadings_PC1, PC2 = loadings_PC2)

# Create separate plots for positive and negative loadings on PC1 and PC2
library(ggplot2)

# Plot for positive loadings on PC1
ggplot(subset(loadings_data, PC1 > 0), aes(x = PC1, y = PC2, label = species)) +
  geom_text() +
  ggtitle("Species with Positive Loadings on PC1")

# Plot for negative loadings on PC1
ggplot(subset(loadings_data, PC1 < 0), aes(x = PC1, y = PC2, label = species)) +
  geom_text() +
  ggtitle("Species with Negative Loadings on PC1")

# Plot for positive loadings on PC2
ggplot(subset(loadings_data, PC2 > 0), aes(x = PC1, y = PC2, label = species)) +
  geom_text() +
  ggtitle("Species with Positive Loadings on PC2")

# Plot for negative loadings on PC2
ggplot(subset(loadings_data, PC2 < 0), aes(x = PC1, y = PC2, label = species)) +
  geom_text() +
  ggtitle("Species with Negative Loadings on PC2")

```




